Spin canonical divisors of log stacky curves

### 2015-06-28 stacky_write_up/spin_cring.tex
Making edits to section 3

\ssec{Inductive Lemmata: Adding Points}
\label{ssec:add-points}
In this section, we prove various inductive statements under the following setup:
Let $\halfcan' \in \di \BQ \otimes X$ be $\BQ$-divisor of $X$ (and often a log half-canonical divisor of some $\sx'$), satisfying certain conditions about $\deg(\lfloor d \halfcan'\rfloor)$ for various values of $d\in\mathbb{N}$ and certain saturation conditions. Let $\halfcan=\halfcan'+P$ for some $P$ possibly in the support of $\halfcan$.  Then we will explicitly describe elements that generate $R_\halfcan$ over $\halfcan$' and initial terms of relations that generate $\initial(I')$ over $\initial(I)$ (allowing the construction of a corresponding Gr\"{o}bner Basis). \todo{Aaron: What new ideas does this description add that wasn't already described in the previous paragraph? Also, in general it's a good idea to be specific about what the most important lemmas are in the section, although I also added this to the general description of section 3.}

First, we introduce some notation from O'Dorney.  \cite{dorney:canonical}
\begin{defn}\label{den:lower-approximation}
If $\frac{\alpha}{\beta}\in \BQ$ (for $\alpha, \beta \in \mathbb{Z}, \beta> 0$ written in reduced form), then a \begin{bf}{best lower approximation}\end{bf} to $\frac{\alpha}{\beta}$ is a a fraction $\frac{c}{d}$ (written in reduced form with $d>0$) such that there is no $\frac{c'}{d'}$ with $0<d'<d$ and $\frac{c}{d}\le \frac{c'}{d'}$.  Moreover, the non-negative best lower approximations of $\frac{\alpha}{\beta}$ form a sequence
\[
	0=\frac{c_0}{d_0} < \frac{c_1}{d_1} < \ldots < \frac{c_r}{d_r} = \frac{\alpha}{\beta}.
\]
\todo{This Moreover statement shouldn't really be part of the definition.}
\end{defn}
\todo{add a picture of lattice}

@@

First, however, we give a criteria to determine if a set of monomials generates the initial ideal of relations, and define notation for a type of monomial ordering. \todo{Perhaps say which parts of the proofs of which future lemmas it will be used in?}

\todo{Somehow, I feel like most of this page can be cut, it seems to mostly vaguely describe what we are doing, and takes up a lot of space.}

\begin{lem}\label{lem:relations_from_generators_induction} 
Suppose $\halfcan, \halfcan' \in \di X \otimes \BQ$ with $\halfcan=\halfcan'+\frac{\alpha}{\beta}P$, that $R_{\halfcan'}$ is generated by $x_1, \ldots, x_m,$
and $R_{\halfcan}$ is minimally generated by $y_1, \ldots, y_n$ over 
$R_{\halfcan'}$ with ideals of relations $I'$ and $I$ respectively \todo{Usually respectively is applied when two things are said at the same time, this doesn't seem to fit the usual gramattical pattern}.  
Suppose there are sets of monomials $S\subseteq R_\halfcan-R_{\halfcan'}$ and $T\subseteq R_\halfcan-(S\cup R_\halfcan')$, and an ordering $\prec$ such 
\begin{enumerate}
\item[(\ref{lem:relations_from_generators_induction}.1)] \todo{It would probably be clearer to just call these conditions 1,2,3 instead of 3.2.1...}  $T\prec S\prec k[x_1, \ldots, x_n]$ (meaning all monomials in $T$ precede all monomials in $S$ which proceed all monomials in $k[x_1, \ldots, x_n]$)
\item[(\ref{lem:relations_from_generators_induction}.2)] $S$ forms a $k$-basis for $R_\halfcan$ over $R_\halfcan'$
\item[(\ref{lem:relations_from_generators_induction}.3)] All monomials in $k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ lie in 
\begin{align*}
	S	\cup \; \langle T\rangle \cup \; \initial_\prec(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]
\end{align*}
\end{enumerate}
Then,
\begin{align*}
	\initial(I) & = \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]
	+ \langle T \rangle.
\end{align*}
\end{lem}
\begin{proof}
Suppose a monomial $f\in T$ has a pole order $a$ at $P$. Then, since $S$ forms a $k$-basis of $R_{\halfcan}$ over $R_{\halfcan'}$, we can write $f = (\sum_{g\in S'} C_g g)+r$ for some finite subset $S'\subseteq S_{\deg(f)}$ and $r\in R_{\halfcan'}$.  This demonstrates that 
\begin{align*}
	\initial(I) & \supseteq \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n] + \langle T \rangle.
\end{align*}
\todo{Aaron: I don't understand why this holds}
Since every monomial of $k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ is an element of either $S$, $\langle T\rangle$, or $\initial_\prec(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]$,
\begin{align*}
	\initial(I) & = \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n] + \langle T \rangle.
\end{align*}
\todo{Aaron: Why can't the initial term lie in S?}
\todo{It might be helpful to say where you use conditions 1 2 and 3.}
\end{proof}

As suggested by the previous lemma, \todo{Reference the lemma, instead of using previous, since things might get moved around} we will need an ordering on monomials, and the following definition will prove useful in the ensuing lemma. \todo{Again, reference this}
\begin{defn}
We equip the graded ring $k[x_1, \ldots x_m]$, with weight $a_i$ and pole order $-\ord_P(x_i)$ attached to $x_i$, \todo{attached to $x_i$ doesn't really sound like the correct terminology to me} the graded $P$ reverse lexicographic order {\bf{g-P-revlex}} \todo{This name seems a bit funny looking to me} if for $f=\prod_{i=1}^m {x_i}^{q_i}\ne g=\prod_{i=1}^m {x_i}^{r_i}$, $f\prec g$ iff \todo{It's better not to use abbreviations like iff in papers. Also, usually in definitions, one uses if in place of iff.} one of the following holds
\begin{enumerate}
\item $\deg(f) < \deg(g)$
\item $\deg(f) = \deg(g)$ and $-\ord(f) > -\ord(g)$ \todo{Order at which point, in this line and the next}
\item $\deg(f) = \deg(g)$, $-\ord(f) = -\ord(g)$, and for $i$ maximal such that $q_i-r_i$ is non-zero, that expression is negative. \todo{I do not understand what this means, maximal under what ordering? Also, what expression are you talking about?}
\end{enumerate}
\end{defn}
We leave it as an exercise to verify that this is an order.  \todo{check for ourselves that this actually is, Aaron: I suggest not to use the phrase ``leave as an exercise'' in a paper.}
We now offer \todo{Maybe state instead of offer?} a lemma that allows us to inductively add $\mathbb{Q}$-points \todo{$\BQ$ points usually refers to maps to $\spec \BQ$. Maybe we can just say a rational multiple of a poitn?} to a divisor and describe the new generators and relations.

@@

\begin{lem}
\label{lem:sat-1}
Let $X$ be a genus $g$ curve and $\halfcan' \in \di X \otimes \BQ$
satisfying $h^0(X, \lfloor{\halfcan'}\rfloor)\ge 1$. Suppose $P$ is not a base-point of $d\halfcan'$ for all $d\in \mathbb{N}^+$ meaning that $R_{\halfcan'}$ has generators (in the necessary degrees \todo{What does this mean?})
$u, x_1, \ldots, x_m$ with $\deg u = 1$, $\ord_P^{\halfcan'}(x_i)=0$ for all $1 \leq i \leq m$, and $\ord_P^{\halfcan'}(u) = 0$.Suppose $u, x_1, \ldots, x_m$ generate $R_{\halfcan'}$ in degree at most $\sigma$ with relations generated
in degree at most $\tau$.  Suppose $\halfcan = \halfcan' + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbb{N}$ such that
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{d\halfcan}\rfloor) = h^0(X,\lfloor d\halfcan'
	\rfloor) + \left\lfloor d\frac{\alpha} {\beta} \right \rfloor \text{ for all } d \in \mathbb{
	N} \text{ such that } d \ge \frac{\beta}{\alpha}.
\end{equation}
\label{eqn:deg1-sat-ind}
Let 
\[
	0 < \frac{c_1}{d_1} < \ldots < \frac{c_n}{d_n} = \frac{\alpha}{\beta}
\]
be the positive best lower approximations of $\frac{\alpha}{\beta}$.
Then,

\begin{enumerate}
\item[(a)] $R_{\halfcan}$ is generated over $R_{\halfcan'}$ by elements $y_1, \ldots, y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$, when viewed as sections of $\so_{\halfcan'}$ \todo{Clearer to say $\deg (y_i) = d_i,\ord_P^{L'} = c_i.$}

\item[(b)] Choose an ordering $\prec$ on $k[u, x_1, \ldots, x_m]$ such that
\[
	\ord_u(f) < \ord_u(g) \implies f\prec g
\]
and equip $k[y_1, \ldots, y_n]$ with g-P-revlex and $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ with a block order \todo{it should be called block order, instead of ``a block order''}.
If $I$ and $I'$ are the ideal of relations of $R_{\halfcan}$ and $R_{\halfcan'}$ respectively, then

\begin{align*}
	\initial_\prec(I) &= \initial_\prec(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] 
										 + \langle U_i: 1 \le i \le n-1 \rangle
										 + \langle V \rangle\\
\end{align*}
where 
$V=\{x_i y_j: 1\le i\le m, 1\le j\le n\}$
 and $U_i$ is the set of monomials of the form $\prod_{j=1}^{i} y_i^{a_i}$ \todo{should this be $y_j^{a_j}$ instead of $y_i^{a_i}$?} with $a_j\in \mathbb{Z}^{\ge 0}, a_i\ne 0$ \todo{this conflation of $j$ and $i$ is confusing me, is $j$ ranging over all indices from $1$ to $i$?} such that
$\sum_{j=1}^i a_j c_j \ge c_{j+1}$
and there does not exist $((b_1, \ldots b_i)\ne (a_1, \ldots a_i)$ such that with all $b_j\le a_j$ and 
$\sum_{j=1}^i b_jc_j\ge c_{j+1})\}$. \todo{what is this close bracket closing?}  \\

@@

\noindent
\todo{why is there no floor in the above equation, as in ~\ref{eqn:deg1_sat_ind_dim}} so in $H^0 (X, \lfloor d\halfcan \rfloor)$ we can find elements with pole
order $i$ at $P$ for any $i \in \{0, \ldots, \lfloor d \frac{\alpha}{
\beta} \rfloor \}$. We will use O'Dorney's method of best lower approximations to
construct the $y_i$'s as desired. \todo{cite where} \todo{What do we desire about the $y_i$?}

>>>>

\ssec{Inductive Lemmata: Adding Points}
\label{ssec:add-points}
%In this section, we prove various inductive statements under the following setup:
%Let $\halfcan' \in \di \BQ \otimes X$ be $\BQ$-divisor of $X$ (and often a log half-canonical divisor of some $\sx'$), satisfying certain conditions about $\deg(\lfloor d \halfcan'\rfloor)$ for various values of $d\in\mathbb{N}$ and certain saturation conditions. Let $\halfcan=\halfcan'+P$ for some $P$ possibly in the support of $\halfcan$.  Then we will explicitly describe elements that generate $R_\halfcan$ over $\halfcan$' and initial terms of relations that generate $\initial(I')$ over $\initial(I)$ (allowing the construction of a corresponding Gr\"{o}bner Basis). \todo{Aaron: What new ideas does this description add that wasn't already described in the previous paragraph? Also, in general it's a good idea to be specific about what the most important lemmas are in the section, although I also added this to the general description of section 3.}

Before proving three inductive lemmas, we introduce some notation from O'Dorney \cite{dorney:canonical}. 
\begin{defn}\label{den:lower-approximation}
If $\frac{\alpha}{\beta}\in \BQ$ (for $\alpha, \beta \in \mathbb{Z}, \beta> 0$ written in reduced form), then a \begin{bf}{best lower approximation}\end{bf} to $\frac{\alpha}{\beta}$ is a a fraction $\frac{c}{d}$ (written in reduced form with $d>0$) such that there is no $\frac{c'}{d'}$ with $0<d'<d$ and $\frac{c}{d}\le \frac{c'}{d'}$. 
\end{defn}
Moreover, the non-negative best lower approximations of $\frac{\alpha}{\beta}$ form a sequence
\[
	0=\frac{c_0}{d_0} < \frac{c_1}{d_1} < \ldots < \frac{c_r}{d_r} = \frac{\alpha}{\beta}.
\]
Refer to figure \ref{fig:s14/5-lattice} for an example of the positive best lower approximations of $\frac{\alpha}{\beta}=\frac{14}{5}$.

\begin{figure}[H]\label{fig:s14/5-lattice}
\includegraphics{pics/spin-lower-approximations-pic-pics.pdf}
\caption{This figure shows the positive best lower approximations shown as $*$ and other lattice points shown in $o$ of $\frac{14}{5}$.  Note that they minimally generate lattice in $\mathbb{N}\times \mathbb{N}$ not above $5y=14x$.}
\end{figure}

@@

Before proving three inductive lemmas, we give a criteria to determine if a set of monomials generates the initial ideal of relations.  \todo{Perhaps say which parts of the proofs of which future lemmas it will be used in?}

\begin{lem}\label{lem:relations_from_generators_induction} 
Suppose $\halfcan, \halfcan' \in \di X \otimes \BQ$ with $\halfcan=\halfcan'+\frac{\alpha}{\beta}P$, that $R_{\halfcan'}$ is generated by $x_1, \ldots, x_m$
and $R_{\halfcan}$ is minimally generated by $y_1, \ldots, y_n$ over 
$R_{\halfcan'}$.  Let $I'$ and $I$ be the ideals relations of $k[x_1, \ldots, x_m]\to R_{\halfcan'}$ and $k[x_1, \ldots, x_m, y_1, \ldots y_n]\to R_{\halfcan}$ respectively.
Suppose there are sets of monomials $S\subseteq R_\halfcan-R_{\halfcan'}$ and $T\subseteq R_\halfcan-(S\cup R_\halfcan')$, and an ordering $\prec$ such 
\begin{enumerate}
\item $S$ forms a $k$-basis for $R_\halfcan$ over $R_\halfcan'$
\item $T\prec S\prec k[x_1, \ldots, x_n]$ (meaning all monomials in $T$ precede all monomials in $S$ which proceed all monomials in $k[x_1, \ldots, x_n]$)
\item All monomials in $k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ lie in 
\begin{align*}
	S	\cup \; \langle T\rangle \cup \; \initial_\prec(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]
\end{align*}
\end{enumerate}
Then,
\begin{align*}
	\initial(I) & = \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]
	+ \langle T \rangle.
\end{align*}
\end{lem}
\begin{proof}
First, notice that $I'\subseteq I$ so 
\[
	\initial(I)\supseteq \initial(I')k[x_1, \ldots, x_m, y_1, \ldots, y_n].
\]

Now, suppose a monomial $f\in T$ has a pole order $a$ at $P$. Then, since $S$ forms a $k$-basis of $R_{\halfcan}$ over $R_{\halfcan'}$ by (1), we can write $f - (\sum_{g\in S'} C_g g)-r=0$ for some finite subset $S'\subseteq S_{\deg(f)}$, $C_g\in k$ for all $g\in S'$, and $r\in R_{\halfcan'}$.  This demonstrates that 
\begin{align*}
	\initial(I) & \supseteq \langle T \rangle.
\end{align*}

Finally, suppose $f\in S$.  Note that by (2) any polynomial $G$ in $x_1,\ldots, x_m, y_1, \ldots, y_n$ with initial term in $S$ cannot have a term in $T$.  Furthermore, since $S$ forms a $k$-basis for $R_\halfcan$ over $R_{\halfcan'}$ by (1), any such polynomial does not map to an element of $R_{\halfcan'}$ (since it has a non-zero constant factor with $f$).  Thus, $G=0$ is not a relation, so $f\not\in \initial(I)$.  Therefore, 
\[
	\initial(I)\subseteq R_\halfcan-S.
\]

Since every monomial of $k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ is an element of either $S$, $\langle T\rangle$, or $\initial_\prec(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]$ by (3),
\begin{align*}
	\initial(I) & = \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n] + \langle T \rangle.
\end{align*}
\end{proof}

As suggested by the the previous lemma, \todo{Reference the lemma, instead of using previous, since things might get moved around} we will need an ordering on monomials, and the following definition will prove useful in the ensuing lemma. \todo{Again, reference this}
\begin{defn}
We equip the graded ring $k[x_1, \ldots x_m]$, with weight $a_i$ and pole order $-\ord_P(x_i)$ attached to $x_i$, \todo{attached to $x_i$ doesn't really sound like the correct terminology to me} the graded $P$ reverse lexicographic order {\bf{g-P-revlex}} \todo{This name seems a bit funny looking to me} if for $f=\prod_{i=1}^m {x_i}^{q_i}\ne g=\prod_{i=1}^m {x_i}^{r_i}$, $f\prec g$ if one of the following holds
\begin{enumerate}
\item $\deg(f) < \deg(g)$
\item $\deg(f) = \deg(g)$ and $-\ord_P(f) > -\ord_O(g)$
\item $\deg(f) = \deg(g)$, $-\ord_P(f) = -\ord_P(g)$, and for $i$ maximal such that $q_i-r_i$ is non-zero, $q_i-r_i$ is negative. \todo{I do not understand what this means, maximal under what ordering?}
\end{enumerate}
\end{defn}
One can easily verify that these satisfy the conditions of an ordering.  
We now state a lemma that allows us to inductively add $\mathbb{Q}$-points \todo{$\BQ$ points usually refers to maps to $\spec \BQ$. Maybe we can just say a rational multiple of a poitn?} to a divisor and describe the new generators and relations.

@@

\begin{lem}
\label{lem:sat-1}
Let $X$ be a genus $g$ curve and $\halfcan' \in \di X \otimes \BQ$
satisfying $h^0(X, \lfloor{\halfcan'}\rfloor)\ge 1$. Suppose $P$ is not a base-point of $d\halfcan'$ for all $d\in \mathbb{N}^+$ meaning that $R_{\halfcan'}$ has generators
$u, x_1, \ldots, x_m$ with $\deg u = 1$, $\ord_P^{\halfcan'}(x_i)=0$ for all $1 \leq i \leq m$, and $\ord_P^{\halfcan'}(u) = 0$.Suppose $u, x_1, \ldots, x_m$ generate $R_{\halfcan'}$ in degree at most $\sigma$ with relations generated
in degree at most $\tau$.  Suppose $\halfcan = \halfcan' + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbb{N}$ such that
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{d\halfcan}\rfloor) = h^0(X,\lfloor d\halfcan'
	\rfloor) + \left\lfloor d\frac{\alpha} {\beta} \right \rfloor \text{ for all } d \in \mathbb{
	N} \text{ such that } d \ge \frac{\beta}{\alpha}.
\end{equation}
\label{eqn:deg1-sat-ind}
Let 
\[
	0 < \frac{c_1}{d_1} < \ldots < \frac{c_n}{d_n} = \frac{\alpha}{\beta}
\]
be the positive best lower approximations of $\frac{\alpha}{\beta}$.
Then,

\begin{enumerate}
\item[(a)] $R_{\halfcan}$ is generated over $R_{\halfcan'}$ by elements $y_1, \ldots, y_n$ such thaty $\deg(y_i)=d_i$ and $-\ord_P(y_i)=c_i$ when viewed as sections of $\so_{\halfcan'}$.

\item[(b)] Choose an ordering $\prec$ on $k[u, x_1, \ldots, x_m]$ such that
\[
	\ord_u(f) < \ord_u(g) \implies f\prec g
\]
and equip $k[y_1, \ldots, y_n]$ with g-P-revlex and $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ with block order \todo{it should be called block order, instead of ``a block order''}.
If $I$ and $I'$ are the ideal of relations of $R_{\halfcan}$ and $R_{\halfcan'}$ respectively, then

\begin{align*}
	\initial_\prec(I) &= \initial_\prec(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] 
										 + \langle U_i: 1 \le i \le n-1 \rangle
										 + \langle V \rangle\\
\end{align*}
where 
$V=\{x_i y_j: 1\le i\le m, 1\le j\le n\}$
 and $U_i$ is the set of monomials of the form $\prod_{j=1}^{i} y_j^{a_j}$ with $a_j\in \mathbb{Z}^{\ge 0}, a_i\ne 0$ such that
\begin{enumerate}
	\item $\sum_{j=1}^i a_j c_j \ge c_{i+1}$\\
	\item there does not exist $((b_1, \ldots b_i)\ne (a_1, \ldots a_i)$ such that with all $b_j\le a_j$ and $\sum_{j=1}^i b_jc_j\ge c_{i+1})$ \\
	\item there does not exist $h<i$ such that $\sum_{j=1}^h y_j^{a_j}\ge c_{h+1}$
\end{enumerate}

@@

\noindent
\todo{why is there no floor in the above equation, as in ~\ref{eqn:deg1_sat_ind_dim}} so in $H^0 (X, \lfloor d\halfcan \rfloor)$ we can find elements with pole
order $i$ at $P$ for any $i \in \{0, \ldots, \lfloor d \frac{\alpha}{
\beta} \rfloor \}$. We will use positive best lower approximations to
construct the generators $\{y_i\}$ as described in the lemma's statement.
